```{r, echo = FALSE, message = FALSE}
library(TeachingSampling)
data(BigLucy)
library(ggplot2)
library(gridExtra)
options(scipen = 100, digits = 2)
set.seed(12345)
library(knitr)
opts_chunk$set(cache = TRUE)
knit_theme$set("acid")
```
# Calibration Estimators

::: {.blockquote}
Calibration [as a process] has become an important methodological instrument in the production of large volumes of statistics. Most statistical agencies have developed software specifically designed to calculate the resulting weights, usually calibrated to auxiliary information available in administrative records and other accurate sources.
@Sar08
:::

The calibration process is the main topic of the most recent articles published on estimation in finite populations and sampling. This phenomenon arises because calibration provides a systematic way to incorporate auxiliary information at the estimation stage of a survey. A calibration estimator is a linear estimator<!-- index: Linear estimator --> with the desirable property of representativeness under any sampling design. Although the term calibration is new, some authors agree that they have used calibration for a long time, before knowing this process by that name.

As @Sar08 states, the most important item in calibration, as a systematic estimation process, is the existence of auxiliary information. If there is no auxiliary information, there is nothing to calibrate to, and therefore there will be no calibration estimators to apply. As will be seen throughout the chapter, general regression estimators can yield the same results as calibration estimators; however, the spirit and essence of their application move in markedly different directions.

But what is a calibration estimator? What is its essence? Below is a brief description of this method:
1. Suppose that an auxiliary information vector $\mathbf{x}_k=(x_{1k}, x_{2k},\ldots,x_{pk})$ of $p$ auxiliary variables is available and known for the individuals selected in the sample.
1. In addition, from administrative records or other reliable sources, the total of the auxiliary information vector $\mathbf{t_X}=\sum_{k\in U}\mathbf{x}_k$ is known.
1. The purpose of the study is to estimate the total of the characteristic of interest using the information given by $\mathbf{x}_k \ \ \ k\in S$.
1. Although the Horvitz-Thompson estimator is unbiased, the estimates are required to satisfy the following restriction:
$$
  \sum_{k\in S}w_k\mathbf{x}_k = \mathbf{t_X}
$$
which is known as the calibration equation.
1. The idea is to seek these weights $w_k$ as close as possible to the inverse of the inclusion probability of the $k$-th element, $d_k=1/\pi_k$.

Although the concept of calibration is new in sampling theory, the essence of the method and the practical spirit of obtaining estimates that match known totals exactly are not new. In fact, this method has been used, and some researchers are still using it, without knowing that it is called calibration. This was the case of @Dem, who approached this topic using contingency tables with internal estimates and known marginal totals. They were the pioneers of the **iterative proportional fitting procedure**, or IPFP.

## IPFP

<!-- index: Iterative proportional fitting procedure -->Suppose that there are two qualitative variables that divide the population into population subgroups. On the one hand, one variable divides the population into $H$ population subgroups, $U_{1 \cdot},\ldots,U_{h \cdot}, \ldots, U_{H \cdot}$; on the other hand, another variable divides the population into $G$ population subgroups, $U_{\cdot1},\ldots,U_{\cdot g}, \ldots, U_{\cdot G}$. As a result, the population is partitioned into $H\times G$ population subgroups, as shown in the following table.

| $U_{11}$ | $\cdots$ | $U_{1g}$ | $\cdots$ | $U_{1G}$ | $U_{1 \cdot}$ |
| --- | --- | --- | --- | --- | --- |
| $\vdots$ |  | $\vdots$ |  | $\vdots$ | $\vdots$ |
| $U_{h1}$ | $\cdots$ | $U_{hg}$ | $\cdots$ | $U_{hG}$ | $U_{h \cdot}$ |
| $\vdots$ |  | $\vdots$ |  | $\vdots$ | $\vdots$ |
| $U_{H1}$ | $\cdots$ | $U_{Hg}$ | $\cdots$ | $U_{HG}$ | $U_{H \cdot}$ |
| $U_{\cdot 1}$ | $\cdots$ | $U_{\cdot g}$ | $\cdots$ | $U_{\cdot G}$ | $U$ |

: Population distribution in the contingency table. {#tbl-t10-static-1}

The sizes of the population subgroups are defined as follows: $N_{hg}=\#U_{hg}$, $N_{h\cdot}=\#U_{h\cdot}$, $N_{\cdot g}=\#U_{\cdot g}$. Note that

$$
N=\sum_{h=1}^H N_{h \cdot} = \sum_{g=1}^G N_{\cdot g}.
$$

In addition, the cell totals of the contingency table follow the following relationship:

| $N_{11}$ | $\cdots$ | $N_{1g}$ | $\cdots$ | $N_{1G}$ | $N_{1 \cdot}$ |
| --- | --- | --- | --- | --- | --- |
| $\vdots$ |  | $\vdots$ |  | $\vdots$ | $\vdots$ |
| $N_{h1}$ | $\cdots$ | $N_{hg}$ | $\cdots$ | $N_{hG}$ | $N_{h \cdot}$ |
| $\vdots$ |  | $\vdots$ |  | $\vdots$ | $\vdots$ |
| $N_{H1}$ | $\cdots$ | $N_{Hg}$ | $\cdots$ | $N_{HG}$ | $N_{H \cdot}$ |
| $N_{\cdot 1}$ | $\cdots$ | $N_{\cdot g}$ | $\cdots$ | $N_{\cdot G}$ | $N$ |

: Distribution of population sizes in the contingency table. {#tbl-t10-static-2}

After data collection and observation in the survey, the final estimate of the totals for each internal cell and marginal cell is available. Thus, $\hat{N}_{hg}$ corresponds to the estimate of $N_{hg}$, $\hat{N}_{h\cdot}$ corresponds to the estimate of $N_{h\cdot}$, $N_{\cdot g}$ corresponds to the estimate of $N_{\cdot g}$, and finally $\hat{N}$ corresponds to the estimate of $N$. In this way, it is possible to use the Horvitz-Thompson estimator by defining

$$
\begin{aligned}
N_{\cdot g}=\sum_{k\in U}z_{hk}  \  \ \ \ \ \
N_{h \cdot}=\sum_{k\in U}z_{gk}.
\end{aligned}
$$

where,

$$
z_{hk}=
 \begin{cases}
 1 &\text{if $k\in U_{h \cdot}$}\\
 0  &\text{otherwise}
 \end{cases} \  \ \ \ \ \
z_{gk}=
 \begin{cases}
 1 &\text{if $k\in U_{\cdot g}$}\\
 0  &\text{otherwise}
 \end{cases}
$$

Using the Horvitz-Thompson estimator guarantees unbiasedness, and the relationship given in the following table is obtained.

| $\hat{N}_{11}$ | $\cdots$ | $\hat{N}_{1g}$ | $\cdots$ | $\hat{N}_{1G}$ | $\hat{N}_{1 \cdot}$ |
| --- | --- | --- | --- | --- | --- |
| $\vdots$ |  | $\vdots$ |  | $\vdots$ | $\vdots$ |
| $\hat{N}_{h1}$ | $\cdots$ | $\hat{N}_{hg}$ | $\cdots$ | $\hat{N}_{hG}$ | $\hat{N}_{h \cdot}$ |
| $\vdots$ |  | $\vdots$ |  | $\vdots$ | $\vdots$ |
| $\hat{N}_{H1}$ | $\cdots$ | $\hat{N}_{Hg}$ | $\cdots$ | $\hat{N}_{HG}$ | $\hat{N}_{H \cdot}$ |
| $\hat{N}_{\cdot 1}$ | $\cdots$ | $\hat{N}_{\cdot g}$ | $\cdots$ | $\hat{N}_{\cdot G}$ | $\hat{N}$ |

: Distribution of estimated population sizes in the contingency table. {#tbl-t10-static-3}

Up to this point, the objective of estimating the internal and marginal cells of the contingency table has been met. However, suppose that, because of administrative records or other reliable sources, it is possible to access the marginal cell totals by both columns and rows. That is, suppose that $N_{\cdot g}$, $g=1,\ldots,G$, and $N_{h \cdot}$, $g=1,\ldots,G$, are known.

Under the preceding assumption, it is possible to construct an algorithm that adjusts the estimates of the internal cells and has the desirable property that, once the algorithm is finished, the row and column sums of the estimates correspond to the known marginal cell totals. This estimation method, based on a very simple algorithm, is known as the **iterative proportional fitting procedure**, or IPFP, and was proposed by @Dem.

### Algorithm

<!-- index: IPFP algorithm -->Although simple and intuitive, the following algorithm is very powerful and has the useful property of converging very quickly if the contingency table has no null values in its internal cells and if the known marginal totals are consistent with the survey implementation.

1. Initialize with
$$
  N_{hg}^{(0)}= \hat{N}_{hg} \ \ \ \ \ \ \ \ g=1,\ldots,G, h=1,\ldots,H
$$
1. For $t=1,2,3,\ldots$
$$
\begin{aligned}
  N_{hg}^{(2t-1)} &=& N_{hg}^{(2t-2)}\frac{N_{h \cdot}}{\sum_{g=1}^GN_{hg}^{(2t-2)}} \ \ \ g=1,\ldots,G, h=1,\ldots,H \\ \\ \\
  N_{hg}^{(2t)}   &=& N_{hg}^{(2t-1)}\frac{N_{\cdot g}}{\sum_{h=1}^HN_{hg}^{(2t-1)}} \ \ \ g=1,\ldots,G, h=1,\ldots,H
\end{aligned}
$$

At first glance, a significant drawback of this method is that it does not take into account the sampling design from which the data used to calibrate to the known auxiliary information arise. However, as will be seen in the following sections, @DS and @Sautor proved that the iterative proportional fitting procedure can indeed be treated as a special case of calibration estimators under the spirit of item 5 in the introduction. The calibration estimators that arise under this frame of reference are known as generalized raking estimators.

### Marco y Lucy

<!-- index: Marco and Lucy -->Returning to our population of firms in the industrial sector, the qualitative variables Level and SPAM are known to form a partition of the population. On the one hand, the variable Level divides the population into three subgroups according to firm characteristics: Large, Medium, and Small. On the other hand, the variable SPAM divides the population into two population subgroups according to advertising strategies, namely `spam_yes` and `spam_no`. In total, the population is divided into $2\times 3= 6$ population subgroups.

Now suppose that a simple random sampling design with sample size $n=400$ has been planned, and that we wish to estimate the total number of firms by industrial group, the total number of firms that use and do not use SPAM, and their corresponding internal nesting in the contingency table, as shown in the following table.

|  | spam_no | spam_yes | Total |
| --- | --- | --- | --- |
| Large | $N_{11}$ | $N_{12}$ | $N_{1 \cdot}$ |
| Medium | $N_{21}$ | $N_{22}$ | $N_{2 \cdot}$ |
| Small | $N_{31}$ | $N_{32}$ | $N_{3 \cdot}$ |
| Total | $N_{\cdot 1}$ | $N_{\cdot 2}$ | $N$ |

: Contingency table for SPAM. {#tbl-t10-static-4}

First, with the help of the `S.SI` function from the\\ `TeachingSampling` package, a probability sample of size $n=2000$ must be selected.

```{r, message = FALSE}
data(BigLucy)
attach(BigLucy)

N <- dim(BigLucy)[1]
n <- 2000
sam <- S.SI(N, n)
sample_data <- BigLucy[sam, ]

attach(sample_data)
head(sample_data)
```

Once the information from each firm selected in the sample has been observed and collected, the `Domains` function from the `TeachingSampling` package is used to obtain two matrices, `spam_no` and `spam_yes`, indicating whether each firm selected in the sample belongs to each of the three levels of the industrial sector.

```{r, message = FALSE}
target_variables <- data.frame(Domains(Level))
domains <- data.frame(Domains(SPAM))
spam_no <- domains[, 1] * target_variables
spam_yes <- domains[, 2] * target_variables
```

The first five elements of the two created matrices are shown below.

```{r, message = FALSE}
head(spam_no)
head(spam_yes)
```

To estimate the marginal totals corresponding to the variables `Level` and `SPAM`, we use the `E.SI` function from the `TeachingSampling` package, applied to the objects `target_variables` and `domains` created in the previous step.

```{r, message = FALSE}
E.SI(N, n, target_variables)
E.SI(N, n, domains)
```

To estimate the internal cells of the contingency table, we use the `E.SI` function from the `TeachingSampling` package, applied to the matrices `spam_no` and `spam_yes` created earlier.

```{r, message = FALSE}
E.SI(N, n, spam_no)
E.SI(N, n, spam_yes)
```

Therefore, the Horvitz-Thompson estimate under simple random sampling is given by @tbl-t10-1. Now suppose that, because of administrative records or other reliable sources, the marginal totals for `Level` and `SPAM` are known: 2905 large firms, 25795 medium firms, and 56596 small firms for the variable `Level`, and 33355 firms that do not use SPAM and 51941 firms that do use SPAM for the variable `SPAM`. It is then possible to use the iterative proportional fitting procedure to calibrate the internal estimates of the contingency table so that they match the known population values exactly. The first step is to create the contingency table in R.

```{r, message = FALSE}
t1 <- as.matrix(E.SI(N, n, spam_no)[1, 2:4])
t2 <- as.matrix(E.SI(N, n, spam_yes)[1, 2:4])
Tab <- data.frame(spam_no = t1, spam_yes = t2)
colnames(Tab) <- c("spam_no", "spam_yes")
```

```{r}
#| label: tbl-t10-1
#| tbl-cap: 'Horvitz-Thompson estimate for the SPAM contingency table.'
colnames(Tab) <- c("spam_no", "spam_yes")
knitr::kable(Tab)
```

Once the contingency table has been created, we implement the algorithm using the `IPFP` function from the `TeachingSampling` package. This function has four arguments. The first argument is `Tab`, which refers to the contingency table resulting from estimation under the probability design. The second argument is `Col`, a vector containing the marginal totals, known population values, of the columns of the contingency table. The third argument is `Row`, a vector containing the marginal totals, known population values, of the rows of the contingency table. Finally, `tol`, which by default is 0.00001, corresponds to the algorithm tolerance. The `IPFP` function returns a contingency table calibrated according to the arguments `Col` and `Tol`. For this particular example, the following output is obtained:

```{r, message = FALSE}
Col <- table(BigLucy$SPAM)
Row <- table(BigLucy$Level)
CalIPFP <- IPFP(Tab, Col, Row, tol = 0.00001)
colnames(CalIPFP)[1:2] <- c("spam_no", "spam_yes")
CalIPFP
```

The following are the comparative tables of the calibrated estimates obtained through the iterative proportional fitting procedure and the corresponding information for the population totals, respectively.

```{r}
#| label: tbl-t10-2
#| tbl-cap: 'Population distribution (unknown) for the SPAM contingency table.'
Tab2 <- table(BigLucy$Level, BigLucy$SPAM)
knitr::kable(Tab2)
```

```{r}
#| label: tbl-t10-3
#| tbl-cap: 'IPFP calibration estimate for the SPAM contingency table.'
colnames(CalIPFP)[1:2] <- c("spam_no", "spam_yes")
knitr::kable(CalIPFP)
```

Note that the relative difference is very small and that the estimates are close to the truth. In these relative terms, this estimate is better than the one induced by the Horvitz-Thompson estimator.

## Theoretical foundations

<!-- index: Calibration estimator -->As established in the previous section, statisticians have tried to use the incorporation of auxiliary information to improve survey estimates. Thus, the regression estimator in all its possible forms requires knowledge of the total of a vector of auxiliary variables. As @DS explain, calibration estimators are a family or class of estimators with a very attractive form, characterized by using calibrated weights that are as close as possible to the original weights or inverses of the inclusion probabilities of the elements selected in the sample; in addition, these calibration estimators respect a set of restrictions, the calibration equations<!-- index: Calibration equation -->.

Consider a finite population $U=\{1,\ldots, k, \ldots, N\}$ from which a probability sample $s$ $(s\subseteq U)$ has been selected, induced by a sampling design $p(\cdot)$. Then $p(s)$ is the probability that sample $s$ was selected. First- and second-order inclusion probabilities are assumed to be strictly positive.

Let $y_k$ be the value of the characteristic of interest for the $k$-th individual in the population, which also has an associated vector of auxiliary values given by $\mathbf{x}_k=(x_{1k}, x_{2k},\ldots,x_{pk})$. Note that $y_k$ and $\mathbf{x}_k$ are observed and known for all elements in the sample. In addition, it is assumed that the population total of the auxiliary information vector $\mathbf{t}_{\mathbf{x}}=\sum_{k\in U}\mathbf{x}_k$ is known through administrative records or other reliable sources.

As in most situations presented in this book, the objective is to estimate the population total of the characteristic of interest, $t_y$. However, the estimator of $t_y$ must be a linear estimator<!-- index: Linear estimator --> of the form

$$
\hat{t}_S(y)=\sum_{k\in S}w_ky_k,
$$

Note that the Horvitz-Thompson estimator takes the preceding form because

$$
\hat{t}_{y,\pi}=\sum_{k\in S}\frac{y_k}{\pi_k}=\sum_{k\in S}d_ky_k,
$$

In addition to linearity, the family of calibration estimators must induce a representative sampling strategy for any sampling design $p(\cdot)$. That is, new weights $w_k$ must be constructed so that they are as close as possible to $d_k=1/\pi_k$ according to some metric, and also satisfy the calibration equations

$$
\sum_{k\in S}w_k\mathbf{x}_k = \mathbf{t_X}.
$$

Note that the weights $w_k$ depend on $S$. For example, under the SRS design, the ratio estimator can be written as

$$
\begin{aligned}
\hat{t}_{yr}=\hat
t_{y\pi}\frac{t_x}{\hat{t}_{x\pi}}&=\sum_{k\in S}\frac{N}{n}\frac{\bar{x}_U}{\bar{x}_S}y_k=\sum_{k\in S}w_ky_k
\end{aligned}
$$

In addition, the weights $w_k$ have the calibration property because

$$
\begin{aligned}
\sum_Sw_kx_k=\sum_S\frac{N}{n}\frac{\bar{x}_U}{\bar{x}_S}x_k=N\frac{\bar{x}_U}{\bar{x}_S}\sum_s\frac{x_k}{n}=N\bar{x}_U=t_x
\end{aligned}
$$

Since there is a variety of estimators that satisfy restriction (10.2.3), weights<!-- index: Calibration weight --> $w_k$ must be found with the following properties [@Este]:
1. **Consistency**: a system of weights or weightings that satisfies (10.2.3) is attractive because it exactly reproduces the known population total for each auxiliary variable<!-- index: Auxiliary variable -->.
1. **Closeness to the basic weights**: the basic weights $d_k=1/\pi_k$ have the attractive property of inducing unbiased estimates with respect to the sampling design used. Any deviation from these weights should be small in order to preserve this property, at least approximately or asymptotically.
1. **Control over the totals of auxiliary variables**: intuition says that the more auxiliary variables used in the calibration process, the better the estimate. This intuitive argument is supported by theory; thus, @Este show that the variance of a calibration estimator decreases as more auxiliary variables are taken into account in calibration.

## Construction

<!-- index: Calibration weight -->To construct these new weights $w_k$, a pseudo-distance<!-- index: Pseudo-distance -->^[A distance function $D(x_1,x_2)$ must satisfy the following properties: i) be strictly positive (nonnegative), meaning $D(x_1,x_2)\geq 0$; ii) $D(x_1,x_2)=0$ only when $x_1=x_2$; iii) be symmetric, that is, $D(x_1,x_2)=D(x_2,x_1)$; and satisfy the triangle inequality, that is, $D(x_1,x_3)\leq D(x_1,x_2)+D(x_2,x_3)$. The function $G(w_k/d_k)$ is a pseudo-distance because it does not necessarily have to satisfy symmetry.] $G(w_k/d_k)$ between $w_k$ and $d_k$ over the whole sample must be minimized. This can be viewed as an optimization problem for the distance over the whole sample given by

$$
\sum_{k\in S} d_k \frac{G(w_k/d_k)}{q_k}
$$

subject to restriction (10.2.3), where $q_k$ $(k\in S)$ form a set of known and strictly positive weights. For the pseudo-distance $G(w_k/d_k)$, it is assumed that

- It must be strictly nonnegative, so that it makes sense as a distance function.
- It must be strictly convex^[A function $G(x)$ is strictly convex if and only if $G(ax_1+(1-a)x_2) < aG(x_1)+ (1-a)G(x_2)$ for all $a\in (0,1)$ and all $x_1\neq x_2$. On the other hand, if the second derivative of $G$ is positive over its entire domain, then $G(x)$ is convex.], so that any local minimum is an absolute minimum.
- $G(1)=0$; that is, the distance between equal weights is zero.
- $G'(1)=0$; when the weights are equal, the function must have a critical point.
- $G?(1)=1$; that critical point must correspond to the minimizer.

In summary, the calibration technique induces a new set of weights $w_k$ that arises from minimizing a pseudo-distance $G(\cdot)$ in the sample, subject to the calibration equations. That is, the new weights must be such that
$$
\sum_{k\in S} w_k\mathbf{x}_k=\sum_U\mathbf{x}_k=\mathbf{t_x}
$$

To solve this optimization problem, we use the Lagrange multiplier technique. Thus, the Lagrange equation is given by the following expression:

$$
\mathcal{L}(w_1,\ldots,w_n,\boldsymbol{\lambda})=\sum_{k\in S} d_k \frac{G(w_k/d_k)}{q_k}-\boldsymbol{\lambda}'\left(\sum_{S}w_k\mathbf{x}_k - \mathbf{t_X}\right)
$$

Differentiating the Lagrange equation with respect to $w_k$ and setting it equal to zero, we have

$$
\frac{\partial \mathcal{L}}{\partial w_k}= \frac{d_k}{q_k}\frac{g(w_k/d_k)}{d_k}-\boldsymbol{\lambda}' \mathbf{x}_k=0
$$

where $g(\omega)=\frac{dG(\omega)}{d\omega}$, and therefore

$$
g(w_k/d_k)=q_k\boldsymbol{\lambda}'\mathbf{x}_k
$$

At this step it is necessary to define a function $F(\cdot)$ such that $F(\cdot)=g^{-1}(\cdot)$; that is, $F(g(\omega))=\omega$. Therefore,

$$
F(g(w_k/d_k))= F(q_k \boldsymbol{\lambda}' \mathbf{x}_k)
$$

which leads us to the value of the new weights:

$$
w_k=d_k F(q_k \boldsymbol{\lambda}' \mathbf{x}_k)
$$

The vector $\boldsymbol{\lambda}$ is obtained by solving the following system of equations:

$$
\sum_{k\in S}\underbrace{d_k F(q_k \boldsymbol{\lambda}' \mathbf{x}_k)}_{w_k}\mathbf{x}_k'=\mathbf{t}_{\mathbf{x}}'
$$

### Distances $G(\cdot)$, $g(\cdot)$, and $F(\cdot)$

<!-- index: Distances -->In general, several types of distances can be used in the construction of a calibration estimator. However, @DS show that all of them asymptotically lead to the same estimator. The most commonly used pseudo-distances are given in Table 10.8. Depending on the choice of each distance, different calibration estimators are obtained. It is also possible to fix two constants $L$ and $U$ and restrict the range of the resulting weights $w_k$ to the interval $(L,U)$. This method is used to avoid extreme or negative weights, which can be eliminated through a good choice of $L$ and $U$.

In summary, the process for obtaining a calibration estimator is as follows:

1. Define a distance $G(\cdot)$ and observe the data $y_k$ and $\mathbf{x}_k$.
1. Solve (10.3.4) for the vector $\boldsymbol{\lambda}$. In some cases, this solution requires iterative procedures.
1. Use $\boldsymbol{\lambda}$ to obtain an estimator of the population total of the characteristic of interest, given by
$$
\hat{t}_{y,cal}=\sum_{k\in S}w_ky_k=\sum_{k\in S}d_k F(q_k \boldsymbol{\lambda}' \mathbf{x}_k)y_k
$$

@DS states that the estimator $\hat{t}_{y,cal}$ will yield estimates close to the unknown population total of the characteristic of interest if there is a strong relationship between $y$ and $\mathbf{x}$. In fact, if $y$ were perfectly explained by $\mathbf{x}$, the variance of the estimator $\hat{t}_{y,cal}$ would be zero for every possible sample.

| Distance | $G(x)$ | $g(x)$ | $F(u)$ |
| --- | --- | --- | --- |
| Chi-square | $\frac{1}{2}(x-1)^2$ | $x-1$ | $1+u$ |
| Entropy | $x\ln(x)-x+1$ | $\ln(x)$ | $\exp(u)$ |
| Hellingster | $2(\sqrt{x}-1)^2$ | $2\left(1-\sqrt{\frac{1}{x}}\right)$ | $(1+\frac{u}{2})^{-2}$ |
| Inverse entropy | $\ln(\frac{1}{x})+x-1$ | $1-\frac{1}{x}$ | $(1+u)^{-1}$ |
| Inverse chi-square | $\frac{1}{2}\frac{(x-1)^2}{x}$ | $\frac{1}{2}\left(1-\frac{1}{x}\right)^2$ | $(1+2u)^{-1/2}$ |

: Examples of pseudo-distances for the calibration process. {#tbl-t10-static-5}

## Some particular cases

<!-- index: Calibration estimator -->@DS examined the statistical properties of $\hat{t}_{y,cal}$ under a series of pseudo-distances $G(\dot)$. This section reviews some particular cases that yield calibration estimators, some known and others new.

### Linear method: chi-square distance

This method, perhaps the most widely used and one of the most important in calibration, is obtained when the chi-square distance is chosen. It calculates, over the whole sample, the distance from the new weights $w_k$ to the classical weights $d_k$ as
$$
\begin{aligned}
\sum_{S}d_kG(w_k/d_k)=\frac{1}{2}\sum_{S}\frac{(w_k-d_k)^2}{d_k}
\end{aligned}
$$

::: {.callout-tip title="Result"}
Under the chi-square distance, and assuming that the weights $q_k=1/c_k$, the calibration estimator takes the form of the general regression estimator.

**Proof.**
From (10.3.3), and using the fact that for this pseudo-distance $F(u)=1+u$, we have

$$
\begin{aligned}
w_k&=d_k F(q_k \boldsymbol{\lambda}' \mathbf{x}_k)\\
&=d_k(1+q_k \boldsymbol{\lambda}' \mathbf{x}_k)\\
&=d_k+d_kq_k \boldsymbol{\lambda}' \mathbf{x}_k
\end{aligned}
$$

and replacing in the calibration equation (10.3.4),

$$
\sum_{s}d_k\mathbf{x}_k'+\sum_{s}d_kq_k \boldsymbol{\lambda}' \mathbf{x}_k\mathbf{x}_k'=\mathbf{t}_{\mathbf{x}}'
$$

Solving conveniently, the Lagrange multiplier is

$$
\boldsymbol{\lambda}'=(\mathbf{t}_{\mathbf{x}}-\mathbf{t_{x\pi}})'\left(\sum_{S}d_kq_k\mathbf{x}_k \mathbf{x}_k'\right)^{-1}
$$

Thus, assuming that $q_k=1/c_k$, we arrive at the calibration estimator for the total of the characteristic of interest, because

$$
\begin{aligned}
w_k&=d_k + d_k(\mathbf{t}_{\mathbf{x}}-\mathbf{t_{x\pi}})'\mathbf{T}^{-1} q_k\mathbf{x}_k
\end{aligned}
$$

where $\mathbf{T}^{-1}$ is defined in (9.2.13). Then,

$$
\begin{aligned}
\hat{t}_{y,cal}&=\sum_{S}w_ky_k\\
&=\sum_s\frac{y_k}{\pi_k}+(\mathbf{t}_{\mathbf{x}}-\hat{\mathbf{t}}_{\mathbf{x}\pi})'\mathbf{T}^{-1}\sum_s\frac{\mathbf{x}_ky_k}{c_k\pi_k}
\end{aligned}
$$
which exactly coincides with expression (9.2.15), which defines the general regression estimator.
:::

The author emphasizes that the general regression estimator is a special case of the family of calibration estimators. It is an error to make assertions about calibration estimators based only on the functional form of the general regression estimator (GREG). Although it is true that a large majority of articles are based on the spirit of the general regression estimator, it must be emphasized that the philosophy of a calibration estimator, while not contradicting the use of the general regression estimator, is quite different from its philosophy. <!-- index: Linear method --><!-- index: Chi-square distance -->

```{r}
#| label: fig-f10-1
#| message: false
#| fig-height: 5
#| echo: false
#| fig-cap: 'Functions $G(x)$ and $F(u)$ using the chi-square distance.'
x <- sort(runif(100, -2, 2))
G <- (1 / 2) * (x - 1)^2
F <- 1 + x

G_x <- ggplot(data = data.frame(x = x, G = G), aes(x = x, y = G)) +
  geom_line() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  labs(title = "G(x)")

F_x <- ggplot(data = data.frame(x = x, G = G), aes(x = x, y = F)) +
  geom_line() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  labs(title = "F(x)")

plot <- gridExtra::grid.arrange(G_x, F_x, nrow = 1)
```

Note that the general regression estimator uses a model to incorporate auxiliary information into the estimation process. As with calibration estimators, not all special cases of the general regression estimator are calibration estimators. The most influential spirit of calibration estimators is not to incorporate a model into the estimation process, but to obtain a set of weights $w_k$. As @Sar08 states, the concepts of *GREG estimation* and *calibration estimation* reflect a clear difference in thinking. The wide variety of possible models generates a broad family of GREG-type estimators. On the other hand, the choice of a distance in the calibration process generates a broad family of calibration estimators, of which the family of linear GREG estimators is a special case.

::: {.callout-tip title="Result"}
Under the chi-square distance, assuming that the weights $q_k=1/x_k$ and that there is only one auxiliary information variable; that is, $\mathbf{x}_k=x_k$, the calibration estimator takes the form of the ratio estimator.

**Proof.**
Under the preceding conditions, we have

$$
\boldsymbol{\lambda}=\frac{\sum_Ux_k}{\sum_Sd_kx_k}-1=\frac{t_x}{\hat{t}_{x,\pi}}-1
$$

Therefore,

$$
w_k=d_k(1+q_kx_k\boldsymbol{\lambda})=d_k(1+\boldsymbol{\lambda})=d_k\left(\frac{t_x}{\hat{t}_{x,\pi}}\right)
$$

Then the calibration estimator takes the following form:

$$
\begin{aligned}
\hat{t}_{y,cal}&=\sum_Sw_ky_k\\
&=\sum_Sd_k\frac{t_x}{\hat{t}_{x,\pi}}y_k\\
&=t_x\frac{\hat{t}_{y,\pi}}{\hat{t}_{x,\pi}}=\hat{t}_{y,r}
\end{aligned}
$$

which coincides with the form of the ratio estimator given by (9.4.15).
:::

### Raking method: entropy distance

The raking method uses the entropy distance as the basis for constructing the calibration estimator. This distance is defined as:

$$
G(x)=x\log(x)-x+1
$$

Note that the distance, over the whole sample, from the new weights $w_k$ to the classical weights $d_k$ is

$$
\sum_Sd_kG(w_k/d_k)=\sum_Sd_k\left( \frac{w_k}{d_k} \ln\left(\frac{w_k}{d_k}\right)-\frac{w_k}{d_k}+1 \right)
$$

From (10.3.3), and using the fact that for this pseudo-distance $F(u) = \exp(u)$, we have

$$
\begin{aligned}
w_k&=d_k F(q_k \boldsymbol{\lambda}' \mathbf{x}_k)\\
&=d_k\exp(q_k \boldsymbol{\lambda}' \mathbf{x}_k)
\end{aligned}
$$

and replacing in the calibration equation (10.3.4),

$$
\sum_{s}d_k\exp(q_k \boldsymbol{\lambda}' \mathbf{x}_k)\mathbf{x}_k'=\mathbf{t}_{\mathbf{x}}'
$$

The preceding system must be solved for $\boldsymbol{\lambda}$, which is a column vector of Lagrange multipliers. After $\boldsymbol{\lambda}$ is determined, the calibrated weights are calculated as $w_k=d_k\exp(q_k \boldsymbol{\lambda}' \mathbf{x}_k)$, and the calibration estimator for the population total of the characteristic of interest is obtained, defined as:

$$
\hat{t}_{y,cal}=\sum_{S}w_ky_k=\sum_{S}d_k\exp(q_k \boldsymbol{\lambda}' \mathbf{x}_k)y_k
$$

What theoretical and practical interpretation would a negative $w_k$ have? A very important aspect of this raking method is that it induces weights $w_k$ that are strictly positive, which does not happen with the linear method. <!-- index: Raking method --><!-- index: Entropy distance -->

```{r}
#| label: fig-f10-2
#| warning: false
#| message: false
#| fig-height: 5
#| echo: false
#| fig-cap: 'Functions $G(x)$ and $F(u)$ using the entropy distance.'
x <- sort(runif(100, -1, 3))
G <- x * log(x) - x + 1
F <- exp(x)

G_x <- ggplot(data = data.frame(x = x, G = G), aes(x = x, y = G)) +
  geom_line() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  labs(title = "G(x)")

F_x <- ggplot(data = data.frame(x = x, G = G), aes(x = x, y = F)) +
  geom_line() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  labs(title = "F(x)")

plot <- gridExtra::grid.arrange(G_x, F_x, nrow = 1)
```

#### Computational aspects for calculating $\boldsymbol{\lambda}$

To calculate the calibration estimator given by (10.4.7), it is necessary to solve the system of equations (10.4.6) for $\boldsymbol{\lambda}$. @DS show that a general solution can be obtained using the iterative Newton-Raphson method. Note that the system of calibration equations<!-- index: Calibration equation --> can be rewritten as a function $\phi$ in terms of $\boldsymbol{\lambda}$, as follows:

$$
\phi(\boldsymbol{\lambda})=\sum_{S}d_k\exp(q_k \boldsymbol{\lambda}' \mathbf{x}_k)\mathbf{x}_k'-\mathbf{t}_{\mathbf{x}}'
$$

Note that the derivative of this function with respect to $\boldsymbol{\lambda}$ is given by:

$$
\phi'(\boldsymbol{\lambda})=\frac{\partial\phi(\boldsymbol{\lambda})}{\partial\boldsymbol{\lambda}}=\sum_{S}d_k\exp(q_k \boldsymbol{\lambda}' \mathbf{x}_k)\mathbf{x}_k'\mathbf{x}_k
$$

for some vector $\boldsymbol{\lambda}$. Then, according to the Newton-Raphson method, a solution would be given by iterating the following expression until convergence:

$$
\boldsymbol{\lambda}^{(a+1)}=\boldsymbol{\lambda}^{(a)}-\left[\phi'\left(\boldsymbol{\lambda}^{(a)}\right)\right]^{-1}\phi(\boldsymbol{\lambda}^{(a)})
$$

Note that the procedure converges when the difference between $\boldsymbol{\lambda}^{(a+1)}$ and $\boldsymbol{\lambda}^{(a)}$ is smaller than a tolerance fixed in advance. In addition, $\boldsymbol{\lambda}^{(0)}=\mathbf{0}$ must be taken into account.

::: {.callout-tip title="Result"}
Under the Newton-Raphson method, the first iteration of the algorithm gives the solution for $\boldsymbol{\lambda}$ when the chi-square distance is used. That is,
$$
\begin{aligned}
\boldsymbol{\lambda}^{(1)}=\hat{\mathbf{T}}^{-1}[\mathbf{t}_{\mathbf{x}}-\mathbf{\hat{t}_{x\pi}}]
\end{aligned}
$$

**Proof.**
$$
\begin{aligned}
\boldsymbol{\lambda}^{(1)}&=\boldsymbol{\lambda}^{(0)}-\left[\phi'\left(\boldsymbol{\lambda}^{(0)}\right)\right]^{-1}\phi(\boldsymbol{\lambda}^{(0)})\\
&=-\left[\sum_{s}d_k\exp(\mathbf{x}_k\boldsymbol{\lambda}^{(0)})\mathbf{x'}_k\mathbf{x}_k\right]^{-1}
\left[\sum_{s}d_k\exp(\mathbf{x}_k\boldsymbol{\lambda}^{(0)})\mathbf{x}_k-\mathbf{t}_x\right]\\
&=-\left[\sum_{s}d_k\mathbf{x'}_k\mathbf{x}_k\right]^{-1}
\left[\sum_{s}d_k\mathbf{x}_k-\mathbf{t}_x\right]\\
&=-\hat{\mathbf{T}}^{-1}[\hat{\mathbf{t}}_{\mathbf{x}\pi}-\mathbf{t}_{\mathbf{x}}]=\hat{\mathbf{T}}^{-1}(\mathbf{t}_{\mathbf{x}}-\hat{\mathbf{t}}_{\mathbf{x}\pi})
\end{aligned}
$$

which coincides with the solution for $\boldsymbol{\lambda}$ given by expression (10.4.2).
:::

From the preceding result, the calibration estimator in the first iteration would be given by

$$
\hat{t}_{y,cal}^{(1)}=\sum_{S}d_k\exp(q_k (\mathbf{t}_{\mathbf{x}}-\hat{\mathbf{t}}_{\mathbf{x}\pi})'\hat{\mathbf{T}}^{-1} \mathbf{x}_k)y_k
$$

#### Programming the estimator with R

<!-- index: R -->This section gives the basic ideas for computational programming of a calibration estimator based on the raking method when a single auxiliary information variable is used. Note that the functions $\phi$ and $\phi'$ are involved in calculating the vector $\boldsymbol{\lambda}$, whose expression is given by equation (10.4.8). Computational programming of this calibration estimator technique can be easily implemented in four simple steps:

1. Program the function $\phi$.
1. Program the function $\phi'$.
1. Use the preceding expressions to calculate the vector $\boldsymbol{\lambda}$.
1. Iterate until convergence.

Four *computational objects* are involved in programming the function $\phi$: the vector $d_k=(1/\pi_1,\ldots,1/\pi_k,\ldots,1/\pi_n)$, the vector $\boldsymbol{\lambda}$, the vector of auxiliary values for each element included in the sample, given by $\mathbf{x}_k=(x_{1k}, x_{2k},\ldots,x_{pk})$, and the vector of population totals of the auxiliary information variables $\mathbf{t}_{\mathbf{x}}$. Thus, the following code creates a function that calculates $\phi$.

```{r, eval = FALSE}
Fi <- function(dk, l, x, tx) {
  e <- matrix(0, n, 1)
  for (k in 1:n) {
    e[k] <- exp(x[k] * l)
  }
  res <- sum(dk * e * x) - tx
  res
}
```

On the other hand, only three *computational objects* are involved in programming the function $\phi'$, and they were also involved in programming $\phi$. The reason is that $\phi'$ is the derivative of $\phi$. These elements are $d_k=(1/\pi_1,\ldots,1/\pi_k,\ldots,1/\pi_n)$, $\boldsymbol{\lambda}$, and $\mathbf{x}_k=(x_{1k}, x_{2k},\ldots,x_{pk})$. The following code creates a function that calculates $\phi'$.

```{r, eval = FALSE}
Fiprima <- function(dk, l, x) {
  e <- matrix(0, n, 1)
  for (k in 1:n) {
    e[k] <- exp(x[k] * l)
  }
  res <- sum(dk * e * x * x)
  res
}
```

At the same time, a function must be created to calculate the calibration estimator. Four *computational objects* are involved in this function: $d_k=(1/\pi_1,\ldots,1/\pi_k,\ldots,1/\pi_n)$, $\boldsymbol{\lambda}$, $\mathbf{x}_k=(x_{1k}, x_{2k},\ldots,x_{pk})$, and finally the vector of values of the characteristic of interest for the sample elements, $y_k=(y_1, y_2,\ldots,y_n)$.

```{r, eval = FALSE}
Cal <- function(dk, l, x, y) {
  w <- matrix(0, n, 1)
  for (k in 1:n) {
    w[k] <- exp(x[k] * l)
  }
  res <- sum(dk * w * y)
  res
}
```

Of course, the preceding code does not work by itself. Note that the previous functions share the computational element $\boldsymbol{\lambda}$; however, this element does not yet exist and must be calculated with iterative methods such as Newton-Raphson. These functions must be assembled by a function that collects them and is able to carry out the final calculation of the calibration estimator.

First, a desired tolerance must be set; in this case, the tolerance is 0.000001. This means that the iterative process stops when $|\boldsymbol{\lambda}^{(a+1)}-\boldsymbol{\lambda}^{(a)}| <- 0.000001$. However, if this condition is not satisfied, the process continues iterating. When the process converges, the functions declared earlier can be used to calculate the value of the estimate.

```{r, eval = FALSE}
tol <- 0.000001
l <- 0
l.k <- 4

while (abs(l - l.k) > tol) {
  l.k <- l - Fi(l, xs, sum(xu)) / Fiprima(l, xs)
  l <- l.k
}

tcal <- Cal(l.k, xs, ys)
```

The preceding programming code is intended as a guide for the student and is not presented as the only computational logic alternative.

Note, however, that although the raking method has the feature that the weights are not negative, unlike what often happens when the linear method is used, they can be highly variable. To solve this problem, @DS propose the logistic and truncated linear methods. These techniques are motivated by the desire to restrict the range of variation of the new calibration weights without excessively altering the calibration estimator. In practice, the statistician wants to avoid extreme weights; the following sections show how these can be eliminated.

### Logistic method

<!-- index: Logistic method -->
```{r}
#| label: fig-f10-3
#| warning: false
#| message: false
#| fig-height: 5
#| echo: false
#| fig-cap: 'Functions $G(x)$ and $F(u)$ using the logistic method with $L=0.4$ and $U=2.5$.'
x <- sort(runif(100, -3, 3))
U <- 2.5
L <- 0.4
A <- (U - L) / ((1 - L) * (U - 1))
G1 <- (x - L) * log((x - L) / (1 - L))
G2 <- (U - x) * log((U - x) / (U - 1))
G <- (1 / A) * (G1 + G2)
F1 <- L * (U - 1) + U * (1 - L) * exp(A * x)
F2 <- (U - 1) + (1 - L) * exp(A * x)
F <- F1 / F2

G_x <- ggplot(data = data.frame(x = x, G = G), aes(x = x, y = G)) +
  geom_line() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  labs(title = "G(x)")

F_x <- ggplot(data = data.frame(x = x, G = G), aes(x = x, y = F)) +
  geom_line() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  labs(title = "F(x)")

plot <- gridExtra::grid.arrange(G_x, F_x, nrow = 1)
```

Commonly known as the Logit $(L,U)$ calibration method, this method fixes two constants $L$ and $U$ such that $L<1<U$. Thus, the following quantity is defined:

$$
A=\frac{(U-L)}{(1-L)(U-1)}
$$

Then the following function is defined:

$$
G(x)=
 \begin{cases}
 \frac{1}{A}\left[(x-L)\lg \frac{x-L}{1-L}+(U-x)\lg  \frac{U-x}{U-1}\right] &\text{if $L<x<U$}\\ \\
 \infty  &\text{otherwise}
 \end{cases}
$$

The corresponding function $F$ is given by

$$
F(u)=
\frac{L(U-1)+U(1-L)\exp(Au)}{U-1+(1-L)\exp(Au)}
$$

The preceding function takes values restricted to the interval $(L,U)$ because $F(-\infty)=L$ and $F(\infty)=U$. Therefore, the new calibration weights are always in the interval $[Ld_k,Ud_k]$.

### Truncated linear method
<!-- index: Truncated linear method -->

To restrict the interval of calibration-weight solutions, it is possible to use the same linear function but restricted to two values $L$ and $U$, such that $L<1<U$. Thus,
$$
G(x)=
 \begin{cases}
 \frac{1}{2}(x-1)^2 &\text{if $L<x<U$}\\ \\
 \infty             &\text{otherwise}
 \end{cases}
$$

In this way, the corresponding function $F$ is given by
$$
F(u)=
 \begin{cases}
 1+u             &\text{if $u\in [L-1,U-1]$} \\ \\
 L               &\text{if $u < L-1$}     \\ \\
 U               &\text{if $u > U-1$}     \\ \\
 \end{cases}
$$

Thus, the new calibration weights are always in the interval $[Ld_k,Ud_k]$.

<!-- index: Truncated linear method -->
```{r}
#| label: fig-f10-4
#| warning: false
#| message: false
#| fig-height: 5
#| echo: false
#| fig-cap: 'Functions $G(x)$ and $F(u)$ using the truncated linear method with $L=0.4$ and $U=2.5$.'
x <- sort(runif(100, -3, 3))
U <- 2.5
L <- 0.4
A <- (U - L) / ((1 - L) * (U - 1))

Gw <- function(x, L, U) {
  n <- length(x)
  Ge <- matrix(0, n, 1)
  for (k in 1:n) {
    if (L < x[k] && x[k] < U) {
      Ge[k] <- (1 / 2) * (x[k] - 1)^2
    } else {
      Ge[k] <- NA
    }
  }
  return(Ge)
}

Fw <- function(x, L, U) {
  n <- length(x)
  Fe <- matrix(0, n, 1)
  for (k in 1:n) {
    if ((L - 1) <= x[k] && x[k] <= (U - 1)) {
      Fe[k] <- 1 + x[k]
    }
    if (x[k] < (L - 1)) {
      Fe[k] <- L
    }
    if (x[k] > (U - 1)) {
      Fe[k] <- U
    }
  }
  return(Fe)
}

G <- Gw(x, L, U)
F <- Fw(x, L, U)

G_x <- ggplot(data = data.frame(x = x, G = G), aes(x = x, y = G)) +
  geom_line() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  labs(title = "G(x)")

F_x <- ggplot(data = data.frame(x = x, G = G), aes(x = x, y = F)) +
  geom_line() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  labs(title = "F(x)")

plot <- gridExtra::grid.arrange(G_x, F_x, nrow = 1)
```

## Calibration and Post-Stratification

<!-- index: Calibration estimator --><!-- index: Post-stratification estimator -->@Sautor first derived the calibration estimator and then explained the post-stratification estimator and the raking estimator, under the IPFP algorithm, as special cases of the calibration method under different distances. This section gives the statistical foundations for constructing these estimators.

### Post-stratification

<!-- index: Post-stratification estimator -->A very important special case of calibration estimators is the complete post-stratification estimator^[The term complete post-stratification is used when the internal totals of the contingency table are known and are used for the calibration process.]. In this case, the number of auxiliary information variables is equal to the number of post-strata that partition the population. This process assumes a partition of the finite population into $G$ groups. Thus $U=(U_1,U_2,\ldots,U_G)$. The characteristic of interest is assumed to be related to $G$ vectors or dummy variables that take the value one if the element belongs to subgroup $U_g$ $(g=1,\ldots,G)$ and zero if the element does not belong to the group. Thus $p=G$, $\mathbf{x}_k=\mathbf{d}_k=(\underbrace{0,0,\ldots, 1, \ldots,0,0}_{G` groups`})'$, and $q_k=1$ for all $k\in U$.

Under the preceding formulation, the vector $\lambda$ takes the following form:

$$
\boldsymbol{\lambda}'=(\lambda_1,\ldots,\lambda_g,\ldots,\lambda_G)
$$

and each entry of the auxiliary information vector for the $k$-th element is given by

$$
x_{kg}=
 \begin{cases}
 1 &\text{if $k\in U_g$}\\ \\
 0 &\text{otherwise}
 \end{cases}
$$

Note that

$$
\mathbf{t_x=}\sum_{k\in U} \mathbf{x}_k'= (N_1,\ldots,N_g,\ldots,N_G),
$$

where $N_g$ corresponds to the total number of elements belonging to population subgroup $U_g$.

::: {.callout-tip title="Result"}
The calibration weights for the post-stratification case are given by

$$
w_k=d_k\frac{N_g }{\hat{N}_{g,\pi}}\ \ \ \ \ \ \ g=1,\ldots,G
$$

and are invariant to the choice of any distance.

**Proof.**
The construction of the calibration estimator for this particular scheme is as follows. First, note that if the $k$-th element belongs to subgroup $U_g$, then

$$
\boldsymbol{\lambda}' \mathbf{x}_k= \lambda_g
$$

Therefore, the calibration restriction given by

$$
\sum_{k\in S}d_k F(\boldsymbol{\lambda}' \mathbf{x}_k)\mathbf{x}_k'=\mathbf{t}_{\mathbf{x}}'
$$

can be rewritten as

$$
\sum_{k\in U_g}d_k F(\lambda_g)=N_g \ \ \ \ \ \ \ g=1,\ldots,G
$$

Therefore, solving the preceding equation, we finally have

$$
F(\lambda_g) =\frac{N_g }{\sum_{k\in U_g}d_k}=\frac{N_g }{\hat{N}_{g,\pi}}\ \ \ \ \ \ \ g=1,\ldots,G
$$

Then, from (10.3.3), the calibration weights are given by

$$
w_k=d_k\frac{N_g }{\hat{N}_{g,\pi}}\ \ \ \ \ \ \ g=1,\ldots,G
$$

Note that the choice of distance did not matter in the construction of the calibration weights.
:::

Therefore, the calibration estimator is given by

$$
\begin{aligned}
\hat{t}_{y,cal}&=\sum_{k\in S}w_ky_k\\
&=\sum_{g=1}^G\sum_{k\in S_g}\frac{N_g }{\hat{N}_{g,\pi}}\frac{y_k}{\pi_k}
\end{aligned}
$$

which is equivalent to the post-stratification estimator.

### Raking

<!-- index: Raking method -->If Deming had known about calibration estimators when the multiplicative distance is used as a frame of reference, he would have been very pleased to see that his method could be generalized and included in statistical science. At first, IPFP was used in a fully pragmatic way: it was simply a matter of making an adjustment so that the internal estimates of the contingency table calibrated to the known totals. Under this frame of reference, IPFP was criticized for being a mathematical rather than statistical method, whose results did not take into account the sampling design used to collect the information. As will be seen in this section, the calibration estimator aimed at estimating internal cells in contingency tables is equivalent to the result of the IPFP method. In fact, the IPFP method is a special case of this scenario, known as raking.

As a particular case, consider the estimation of a two-way contingency table with calibration on the marginal totals. Therefore, the population partition follows the pattern in the following table.

| $U_{11}$ | $\cdots$ | $U_{1g}$ | $\cdots$ | $U_{1G}$ | $U_{1 \cdot}$ |
| --- | --- | --- | --- | --- | --- |
| $\vdots$ |  | $\vdots$ |  | $\vdots$ | $\vdots$ |
| $U_{h1}$ | $\cdots$ | $U_{hg}$ | $\cdots$ | $U_{hG}$ | $U_{h \cdot}$ |
| $\vdots$ |  | $\vdots$ |  | $\vdots$ | $\vdots$ |
| $U_{H1}$ | $\cdots$ | $U_{Hg}$ | $\cdots$ | $U_{HG}$ | $U_{H \cdot}$ |
| $U_{\cdot 1}$ | $\cdots$ | $U_{\cdot g}$ | $\cdots$ | $U_{\cdot G}$ | $U$ |

: Population partition. {#tbl-t10-static-6}

Assume that $q_k=1$ for all $k\in U$ and $\mathbf{x}_k=(\mathbf{d}_{1k}',\mathbf{d}_{1k}')$, where $\mathbf{d}_{1k}$ is a vector of $H$ dummy variables denoting which post-stratum the $k$-th element belongs to, and $\mathbf{d}_{2k}$ is a vector of $G$ dummy variables denoting which post-stratum the $k$-th element belongs to. Note that

$$
\mathbf{t_x=}\sum_{k\in U} \mathbf{x}_k'= (N_{1\cdot},\ldots,N_{h\cdot},\ldots,N_{H\cdot},N_{\cdot1},\ldots,N_{\cdot g},\ldots,N_{\cdot G})
$$

Let $\mathbf{u}=(u_1,\ldots,u_H)'$ be a vector of order $H$ and $\mathbf{v}=(v_1,\ldots,v_G)'$ a vector of order $G$. Defining $\boldsymbol{\lambda}'=(\mathbf{u}',\mathbf{v}')$, if the $k$-th element belongs to cell $U_{hg}$, then

$$
F(q_k\boldsymbol{\lambda}'\mathbf{x}_k)=F(u_h+v_g)
$$

Therefore, the calibration equations<!-- index: Calibration equation --> (10.5.6) can be written as the following system of equations:

$$
\begin{aligned}
\sum_{h=1}^H \hat{N}_{hg,\pi}F(u_h+v_g)=N_{\cdot g} \ \ \ \ \ \ \ g=1,\ldots,G \\
\sum_{g=1}^G \hat{N}_{hg,\pi}F(u_h+v_g)=N_{h \cdot} \ \ \ \ \ \ \ h=1,\ldots,H
\end{aligned}
$$

where $\hat{N}_{hg,\pi}$ corresponds to the Horvitz-Thompson estimator of $N_{hg}$. If the entropy distance is used, then

$$
F(u_h+v_g)=\exp(u_h+v_g)=\exp(u_h)\exp(v_g)
$$

Therefore, the system of equations given by (10.5.12) and (10.5.13) takes the following form:

$$
\begin{aligned}
\exp(u_h)=\dfrac{N_{\cdot g}}{\sum_{g=1}^G \hat{N}_{hg}\exp(v_g)} \ \ \ \ \ \ \ h=1,\ldots,H \\
\exp(v_g)=\dfrac{N_{h \cdot}}{\sum_{h=1}^H \hat{N}_{hg}\exp(v_h)} \ \ \ \ \ \ \ g=1,\ldots,G
\end{aligned}
$$

A solution to the preceding system of equations is obtained by iterating the IPFP algorithm until convergence, as follows.

1. Set $\exp(v_g)=1$ and calculate $\exp(u_h)$ in (10.5.15).
1. Then insert this value of $\exp(u_h)$ in (10.5.16) and calculate a new value of $\exp(v_g)$.
1. Iterate until convergence.

After the algorithm has finished, the calibration estimator for the total of cell $U_{hg}$ is given by

$$
\hat{N}_{hg,cal}=\hat{N}_{hg,\pi}\exp(u_h+v_g)=\hat{N}_{hg,\pi}\exp(u_h)\exp(v_g)
$$

and the new calibrated weights are $w_k=d_k\exp(u_h+v_g)=d_k\exp(u_h)\exp(v_g)$ if the $k$-th element belongs to cell $U_{hg}$.

## Variance of calibration estimators

<!-- index: Variance of calibration estimators -->We close this chapter with an important property of calibration estimators.

::: {.callout-tip title="Result"}
The calibration estimator is asymptotically equivalent to the general regression estimator under the following regularity conditions:
1. $\lim \dfrac{\mathbf{t}_{\mathbf{x}}}{N}$ exists
1. $\dfrac{\hat{\mathbf{t}}_{\mathbf{x}\pi}-\mathbf{t}_{\mathbf{x}}}{N}\rightarrow \mathbf{0}$ in probability^[The frame of reference for this probability measure is given by the sampling design used in the sampling strategy.]
1. $\sqrt{n}\dfrac{\hat{\mathbf{t}}_{\mathbf{x}\pi}-\mathbf{t}_{\mathbf{x}}}{N}$ converges in distribution to the multivariate normal $N(\mathbf{0},\mathbf{A})$

**Proof.**
The proof of the preceding result is beyond the scope of this book. However, interested readers may consult @DS.
:::

In particular, under the preceding result, the calibration estimator shares the same asymptotic expectations and the same asymptotic variances as the general regression estimator. This can be seen through the following heuristic argument:

- It is assumed that for large sample sizes, the Horvitz-Thompson estimator, $\hat{\mathbf{t}}_{\mathbf{x}\pi}$, is close to the population total of the auxiliary information characteristics, $\mathbf{t}_{\mathbf{x}}$. This holds because $\hat{\mathbf{t}}_{\mathbf{x}\pi}$ is a consistent estimator for $\mathbf{t}_{\mathbf{x}}$.
- Then, following equation (10.3.4), the value of $F(\cdot)$ should be close to one and the value of $\boldsymbol{\lambda}$ should be close to $\mathbf{0}$.
- However, by the construction of the functions $F(\cdot)$ and because $F(0)=F'(0)=1$, all functions $F(\cdot)$ should behave the same way in a neighborhood of $0$.
- Therefore, all functions $F(\cdot)$ can be approximated by the function $F(u)=u+1$.
- That is, the same function that corresponds to the general regression estimator.

::: {.callout-tip title="Result"}
The approximate variance and the variance estimate of the calibration estimator are given by
$$
AV(\hat{t}_{y,cal})=\sum\sum_U\Delta_{kl}(d_kE_k)(d_lE_l)
$$
$$
\widehat{Var}(\hat{t}_{y,cal})=\sum\sum_S\frac{\Delta_{kl}}{\pi_{kl}}(w_ke_k)(w_le_l)
$$
respectively, where $E_k=y_k-\mathbf{x}_k'\mathbf{B}$ and $\mathbf{B}$ satisfies the normal equations in the construction of the regression estimator. Also, $e_k=y_k-\mathbf{x}_k'\hat{\mathbf{B}}$ and $\hat{\mathbf{B}}$ is an estimator of $\mathbf{B}$.

:::
## Marco y Lucy

<!-- index: Marco and Lucy -->Returning to the practical estimation exercise, suppose that the government wants to obtain an estimate of the total taxes contributed by the industrial sector in the last fiscal year. These estimates must be very precise because part of the national budget will be redesigned based on these results.

On this occasion, the government provides the statistician with a sampling frame that includes the identification and location of all firms belonging to the industrial sector. In addition, the government has knowledge of the population total of two auxiliary information characteristics: the population total of the variable `Employees`, equal to 151950; the population total of the variable `Income`, equal to 1035217; and, of course, the population total of the number of firms in the industrial sector, equal to 2396.

Under the preceding scheme, several sampling strategies were planned, using a simple random design of 400 firms and calibration estimators under several distances. The following computational code was used to select this sample.

```{r, warning = FALSE, message = FALSE}
data(BigLucy)
attach(BigLucy)
N <- dim(BigLucy)[1]
n <- 2000
sam <- sample(N, n)
sample_data <- BigLucy[sam, ]
attach(sample_data)
```

Once the sample was selected, the R `sampling` package was used to calculate the calibration estimators. In particular, the `calib` function was used to calculate the weights $w_k$ of the calibration estimator. This function has several arguments, including the following: `Xs`, the matrix containing the values of the auxiliary information characteristics for the individuals included in the sample; `d`, corresponding to the inverse of the inclusion-probability weights of the sample elements; `tx`, which corresponds to the population total of the calibration variables; and `method`, which includes four possible distances: the chi-square distance, whose option in `calib` is `method="linear"`; the entropy distance, whose option in `calib` is `method="raking"`; and the logistic and truncated methods, whose options in `calib` are `method="logit"` and `method="truncated"`, respectively.

Calibration estimates were calculated using the four methods, and the code used is shown below.

```{r}
library(sampling)
tx1 <- sum(BigLucy$Income)
tx2 <- sum(BigLucy$Employees)

ys <- data.frame(Income, Employees, Taxes)
Xs <- cbind(1, Income, Employees)
piks <- rep(n / N, times = n)
tx <- c(N, tx1, tx2)

w1 <- calib(Xs, d = 1 / piks, tx, method = "linear")
w2 <- calib(Xs, d = 1 / piks, tx, method = "raking")
w3 <- calib(Xs, d = 1 / piks, tx, method = "logit", bounds = c(0.75, 1.2))
w4 <- calib(Xs, d = 1 / piks, tx, method = "truncated", bounds = c(0.75, 1.2))
```

The `calib` function only calculates the weights involved in the calibration equations. To calculate the final estimate of the total of the characteristic of interest `Taxes`, the relevant quantities must be multiplied. Thus, the following code was used to calculate the four estimates.

```{r}
tcal1 <- t(w1 / piks) %*% as.matrix(ys)
tcal1
tcal2 <- t(w2 / piks) %*% as.matrix(ys)
tcal2
tcal3 <- t(w3 / piks) %*% as.matrix(ys)
tcal3
tcal4 <- t(w4 / piks) %*% as.matrix(ys)
tcal4
```

## Discussion

<!-- index: Calibration estimator -->@Sar08 states that the definition of the calibration approach for estimating totals in finite populations follows these processes:

1. Calculate new weights that incorporate specific auxiliary information and are restricted by the calibration equation.
1. Use these new weights to construct linear estimators.
1. Obtain approximately unbiased estimates in the presence of nonresponse<!-- index: Nonresponse --> and other nonsampling errors.

At the same time, @Sar08 concludes that there are six ideas worth exploring further. The following criteria are those some statisticians have used to emphasize the practical use of calibration estimators:

- **As a linear weighting method:** calibration has an intimate link with practice. The attachment of official statistics agencies to weighting methods is a powerful habit in practice that began with weighting units by the inverse of their inclusion probability and continued with the weights arising from the post-stratification approach. Calibration weights extend these earlier ideas. Calibration is new as a term in sampling, nearly 15 years old, but not new as a technique for producing weights. For example, quota sampling<!-- index: Quota sampling --> is a form of nonprobability sampling that induces estimates calibrated to the demographic totals of the study population. Weighting the observed values of the characteristics of interest was a very important topic before the term calibration became popular. Some authors derived these weights with the argument that they should differ as little as possible from the original weights. Other authors found the weights by recognizing that a linear regression estimator could be written as a weighted sum of the values of the characteristic of interest. From there arose terms such as sampling weight, regression weight, and case weight.
- **As a systematic way to use auxiliary information:** calibration provides a systematic way to involve auxiliary information. In most practical applications, calibration provides a simple approach to incorporating this information into the estimation stage. Auxiliary information was used to improve the precision of estimates long before the term calibration became popular. Hundreds of articles were written with this purpose in mind. Today, calibration offers a path for incorporating this auxiliary information. For example, calibration can be used effectively in surveys where auxiliary information is available at different levels. In two-stage sampling, auxiliary information may exist for the first-stage units, the clusters, and other information may exist for the second-stage units, elements or clusters.
- **As an approach to achieving consistency:** sometimes the term calibration refers to a way of obtaining consistent estimates^[In this section, consistent is used in the sense of consistency with the totals of the auxiliary information.]. The calibration equations<!-- index: Calibration equation --> impose consistency on the vector of weights; thus, when it is applied to the auxiliary variables, the result will be consistent with the totals of these variables. A desire to promote credibility in official statistics is one reason why agencies seek consistency. When the primary motivation for calibration is not agreement with the totals of the auxiliary information but rather reducing variance and bias due to nonresponse, the vector of weights is said to be balanced.
- **As a reason for transparency and convenience:** the calibration approach has gained popularity in real applications because the resulting estimates are easy to interpret and motivate, since they are directly related to the weights induced by the sampling design. Calibration on known totals gives the user a natural and transparent form of estimation. Users who understand sampling weights appreciate the calibration method because it subtly modifies the original weights while also respecting the totals of the auxiliary information and keeping bias negligible. There is another advantage appreciated by users: in most applications, calibration induces a single vector of weights applicable to all variables involved in the study. This last reason makes this method highly valued in official agencies that handle very large surveys.
- **In combination with other terms:** some authors use the word calibration in combination with other terms to describe various lines of thought. Among this proliferation of terms are model calibration, G calibration, harmonized calibration, calibration at a higher level, regression calibration, nonlinear calibration, super-generalized calibration, neural-network model calibration, and calibration based on local polynomial models, among others. Calibration plays a significant role in indirect sampling methods (see Chapter 12). The term has also been used, although in a different spirit, in concepts such as calibrated imputation and biased calibration.
- **As a new direction of thought:** if calibration represents a new approach clearly demarcated from its predecessors, then it is time to ask: Does calibration generalize previous theories? Does calibration provide better answers to important questions than previous estimation approaches? In practice, the statistician encounters details such as nonresponse, sampling frame deficiencies, and measurement errors. It is true that processes such as imputation and reweighting for nonresponse are widely disseminated and used in practice. However, there remains an unsatisfying feeling when using these methods because they are not framed within an exhaustive theory of inference in finite populations. Most theoretical articles deal with parameter estimation under an ideal world, which does not exist in practice, where nonresponse and other nonsampling errors are absent.

## Optimal calibration estimators

<!-- index: Optimal calibration estimator -->As @Wu1 states, there are two variants in the construction of a calibration estimator: one is given by the choice of distance and the other by the set of calibration equations^[Note that if the auxiliary information vector has $P$ auxiliary variables, then there will be $P$ calibration equations.]. In areas such as demography, it is customary to calibrate on many variables in order to estimate the known totals of the auxiliary variables with zero variance, regardless of whether the resulting estimator may lose efficiency. In these terms, it would be better to use the fewest possible calibration equations so as not to spoil the good behavior of the estimator. The question the researcher must ask is: what is the best calibration equation to use in constructing an estimator of this type?

If $u_k=u(\mathbf{x}_k)$, where $u(\cdot)$ is a real-valued function, then a new way to construct a calibration estimator is given by obtaining weights $w_k$ restricted^[Under this frame of reference, there is a reduction in the number of restrictions used in calibration.] to
$$
\sum_{k\in S}w_ku(\mathbf{x}_k)=\sum_{k\in U}u(\mathbf{x}_k)
$$ {#eq-caloptu}
Therefore, the question becomes clearer: which function $u(\cdot)$ makes the estimator $\hat{t}_{ycal}$ more efficient? It is well known that under sampling-design-based inference, there is no uniformly minimum-variance unbiased estimator [@CSW]. However, it is possible to obtain an optimal estimator under inference assisted by superpopulation models. The answer to these questions is given by the proposal of @Wu1, who constructed an optimal calibration estimator assuming that the responses $y_k$ can be viewed as realizations of the following semiparametric superpopulation model:
$$
E_\xi(y_k|\mathbf{x}_k)=\mu(\mathbf{x}_k,\boldsymbol{\theta}),\ \ \
Var_\xi(y_k|\mathbf{x}_k)=[\nu(\mathbf{x}_k)]^2\sigma^2 \ \ \ \ \ ,
$$ {#eq-modeloopt}

where $\mu(\cdot\ , \cdot)$ and $v(\cdot)$ are known functions, and $\boldsymbol{\theta}$ and $\sigma^2$ are unknown model parameters. It is assumed that the $y_k$, $k\in U$, are conditionally independent given the $\mathbf{x}_k$. Note that $\nu$ may be a known function of $\mu$, as in generalized linear models.

Optimal estimators, assisted by a superpopulation model $\xi$, that minimize the expected value of the sampling-design-based variance, $E_\xi(Var_p(\hat{Y}))$, have been discussed^[The terms $E_p$ and $Var_p$ refer to expectation and variance under a sampling design $p(\cdot)$, and $E_\xi$ and $Var_\xi$ denote expectation and variance under a superpopulation model $\xi$.] by many authors. For example, in @Isaki this expected variance was called anticipated variance.

::: {.callout-tip title="Theorem 1 of @Wu1"}
Let $t_{y,C_u}$ be a calibration estimator of the population total of the characteristic of interest, constructed using restriction (@eq-caloptu), where $C_u=\left\{ u(\mathbf{x}_1),u(\mathbf{x}_2),\ldots,u(\mathbf{x}_N)\right\}$ is the family of vectors of all possible real-valued functions applied to the auxiliary information. Within the class of calibration estimators $t_{y,C_u}$, the choice of
$$
C_{\mu}=\left\{\mu(\mathbf{x}_1,\boldsymbol{\theta}),\mu(\mathbf{x}_2,\boldsymbol{\theta}),\ldots,\mu(\mathbf{x}_N,\boldsymbol{\theta})\right\}
$$
minimizes $E_\xi(Var_p(\hat{Y}))$ under the superpopulation model given by (@eq-modeloopt) and assuming regularity conditions in the sampling design.

:::
With this result, we can proceed to the construction of the optimal calibration estimator resulting from minimizing chi-square subject to the following restriction:
$$
\sum_{k\in S}w_k\hat{\mu}_k=\sum_{k\in U}\hat{\mu}_k
$$

where $\hat{\mu}_k=\mu(\mathbf{x}_k,\hat{\boldsymbol{\theta}})$. The reason is that the values of the vector $\boldsymbol{\theta}$ are unknown and must be replaced by an estimator based on the selected sample, given by $\hat{\boldsymbol{\theta}}$. The minimization is carried out using a Lagrange multiplier, as in @Dev. In this way, it is very easy to obtain the expression of the optimal calibration estimator, which is given by [@Wu3]
$$
\begin{aligned}
\hat{t}_{y,opt}&=\sum_{k\in S}w_ky_k\\
&=\hat{t}_{y\pi}+(t_{\hat{\mu}}-\hat{t}_{\hat{\mu}\pi}){\hat{B}_y}
\end{aligned}
$$

where $t_{\hat{\mu}}=\sum_{k\in U}\hat{\mu}_k$ is the population total of the functions $\hat{\mu}$, and $\hat{t}_{\hat{\mu}\pi}$ is its corresponding Horvitz-Thompson estimator, with
$$
\hat{B}_y=\frac{\sum_{k\in S}d_kq_k\hat{\mu}_ky_k}{\sum_{k\in S}d_kq_k\hat{\mu}_k^2}
$$

In summary, optimal calibration estimators have been studied in depth in @Wu3 and @Wu1, and their foundation lies in model-assisted inference. To motivate the optimality conditions, a general semiparametric superpopulation model given by (@eq-modeloopt) was used. These calibration estimators for the population total of the characteristic of interest have the following characteristics:
1. A chi-square distance whose weight factors satisfy $q_k>0$ and are also such that $N^{-1}\sum_{k=1}^N q_k^2=O(1)$.
1. A single restriction, given by a dimension reduction $u_k=\mu(\mathbf{x}_k,\boldsymbol{\theta})$, where the functional form $\mu(\cdot\ ,\cdot)$ may be arbitrary.

Some of the most important results of this method can be summarized as follows [@Wu1]:

- Let $\hat{\boldsymbol{\theta}}=(\sum_{k\in S}d_kq_k\mathbf{x}_k\mathbf{x}_k')^{-1} \sum_{k\in S}d_kq_k\mathbf{x}_ky_k$. If $u_k=\mathbf{x}_k'\boldsymbol{\theta}$ is used as the calibration variable, the resulting calibration estimator is identical to the conventional calibration estimator given by $\hat{t}_{ycal}$. Therefore, the class of estimators resulting from this method is very general because it includes the original estimator as a special case.
- For any consistent estimator of $\boldsymbol{\theta}$ such that $\hat{\boldsymbol{\theta}}=\boldsymbol{\theta} + o_p(1)$, if $\boldsymbol{\theta}$ is replaced by $\hat{\boldsymbol{\theta}}$ in the calibration equations, the resulting calibration estimator does not change asymptotically.
- Optimal calibration estimators obtained using $u_k=E_\xi(y_k \mid \mathbf{x}_k)=\mu(\mathbf{x}_i,\boldsymbol{\theta})$ are optimal under the minimum expected variance criterion.
- Optimal calibration estimators are optimal under the superpopulation model $\xi$, but even if the considered model is incorrectly specified, these estimators remain consistent.

Since there is no uniformly minimum-variance unbiased estimator, the only choice of $u(\cdot)$ that makes $\hat{t}_{yopt}$ an estimator with the preceding characteristics is $u(\mathbf{x}_k)=y_i$, which is of course practically useless. Therefore, one should make $u(\mathbf{x}_k)\approx y_k$.

The reader should note that the structure of the model $\xi$ given by (@eq-modeloopt) is very general and includes two important cases. The first is the linear or nonlinear regression model given by
$$
y_k=\mu(\mathbf{x}_k,\boldsymbol{\theta})+\nu_k\varepsilon_k
$$

where the $\varepsilon_k$ are independent and identically distributed random variables with $E_\xi(\varepsilon_k)=0$, $Var_\xi(\varepsilon_k)=\sigma^{2}$, and $\nu_k=\nu(\mathbf{x}_k)$ is a known and strictly positive function.

The second case refers to the generalized linear model given by
$$
g(\mu_i)=\mathbf{x}'_k\boldsymbol{\theta}, \ \ \ Var_\xi(y_k|\mathbf{x}_k)=\nu(\mu_k)
$$ {#eq-glmopt}

where $\mu_k=E_\xi(y_k|\mathbf{x}_k)$, $g(\cdot)$ is a link function, and $\nu(\cdot)$ is a variance function.

The behavior of optimal calibration estimators under a linear model and a log-linear model is described below.

```{r}
#| label: fig-f10-5
#| warning: false
#| message: false
#| fig-height: 5
#| echo: false
#| fig-cap: 'Linear behavior of the characteristic of interest explained by the auxiliary information.'
x <- runif(1000, 1, 30)
y <- rnorm(1000, 1.2 * x, sqrt(x))
dat <- data.frame(x = x, y = y)

ggplot(dat, aes(x = x, y = y)) +
  geom_point(shape = 1) +
  geom_smooth()
```

If the auxiliary information explains the characteristic of interest linearly, as observed in Figure 10.5, then the argument expressed in @DS would make sense; motivated by the ratio estimator, they argue that \textsl{?...the [calibration] weights that fit the auxiliary variables well [exactly reproduce their population total] also fit the study variable well...?}

In the multivariate case, the function that makes the calibration estimator optimal is given by
$$
u(\mathbf{x}_k,\boldsymbol{\theta})=\mathbf{x}_k'\boldsymbol{\theta}=\theta_0+\theta_1x_{k1}+...+\theta_Px_{kP}
$$

where $\boldsymbol{\theta}=(\theta_0,\theta_1,...,\theta_P)$ is estimated through weighted least squares, as in multiple regression. Therefore, the characteristic of interest follows the following superpopulation model:

$$
y_k=\mathbf{x}_k'\boldsymbol{\theta}+\nu_k\varepsilon_k
$$

where the $\varepsilon_k$ are independent and identically distributed with $E_\xi(\varepsilon_k)=0$ and $Var_\xi(\varepsilon_k)=\sigma^{2}$, and $\nu_k=\nu(\mathbf{x}_k)=1$. Therefore, when estimating $\boldsymbol{\theta}$ using the least squares technique, we have
$$
\begin{aligned}
\hat{\boldsymbol{\theta}}&=\left(\sum_{k\in S}q_kd_k\mathbf{x}_k\mathbf{x}_k'\right)^{-1}\sum_{k\in S}q_kd_k\mathbf{x}_ky_k\\
&=(\mathbf{X}'\mathbf{V}^{-1}\mathbf{X})^{-1}\mathbf{X}'\mathbf{V}^{-1}\mathbf{y}
\end{aligned}
$$

where $\mathbf{V}=diag(d_1q_1,\ldots,d_nq_n)=\frac{1}{\sigma^2}diag(d_1,\ldots,d_n)$.

::: {.callout-tip title="Result"}
In this way, the calibration estimator of the population total resulting from the preceding superpopulation model is given by
$$
\hat{t}_{y,opt}=t_{y\pi}+(\mathbf{t}_{\mathbf{x}}-\hat{\mathbf{t}}_{\mathbf{x}\pi})'\hat{\boldsymbol{\theta}}
$$

**Proof.**
$$
\begin{aligned}
\hat{t}_{y,opt}&=\hat{t}_{y\pi}+(t_{\hat{\mu}}-\hat{t}_{\hat{\mu}\pi}){\hat{B}_y}\\
&=\hat{t}_{y\pi}+(\sum_{k\in U}\hat{\mu_k}-\sum_{k\in U}d_k\hat{\mu}_k){\hat{B}_y}\\
&=\hat{t}_{y\pi}+(\sum_{k\in U}\mathbf{x}_k'\hat{\boldsymbol{\theta}}-\sum_{k\in U}d_k\mathbf{x}_k'\hat{\boldsymbol{\theta}}){\hat{B}_y}\\
&=\hat{t}_{y\pi}+(\sum_{k\in U}\mathbf{x}_k'-\sum_{k\in U}d_k\mathbf{x}_k')\hat{\boldsymbol{\theta}}{\hat{B}_y}\\
&=\hat{t}_{y\pi}+(\sum_{k\in U}\mathbf{x}_k-\sum_{k\in U}d_k\mathbf{x}_k)'\hat{\boldsymbol{\theta}}{\hat{B}_y}\\
&=\hat{t}_{y\pi}+(\sum_{k\in U}\mathbf{x}_k-\sum_{k\in U}d_k\mathbf{x}_k)'\hat{\boldsymbol{\theta}}
\end{aligned}
$$
because ${\hat{B}_y}=1$. This follows from the definition of $\hat{B}_y$, taking into account that
$$
\hat{\mu}_k=\mathbf{x}_k'\boldsymbol{\theta}=x_k'(\mathbf{X}'\mathbf{V}^{-1}\mathbf{X})^{-1}\mathbf{X}'\mathbf{V}^{-1}\mathbf{y}
$$
Therefore,
$$
\begin{aligned}
\sum_{k\in S}d_kq_k\hat{\mu}_k^2&=\mathbf{y}'\mathbf{V}^{-1}\mathbf{X}
(\mathbf{X}'\mathbf{V}^{-1}\mathbf{X})^{-1}\mathbf{X}'\mathbf{V}^{-1}\mathbf{X}
(\mathbf{X}'\mathbf{V}^{-1}\mathbf{X})^{-1}\mathbf{X}'\mathbf{V}^{-1}\mathbf{y}\\
&=\mathbf{y}'\mathbf{V}^{-1}\mathbf{X}(\mathbf{X}'\mathbf{V}^{-1}\mathbf{X})^{-1}\mathbf{X}'\mathbf{V}^{-1}\mathbf{y}\\
&=\sum_{k\in S}d_kq_k\hat{\mu}_ky_k
\end{aligned}
$$

:::

Note that the term $\hat{B}_Y$ is equal to one and therefore disappears, making the optimal calibration estimator identical to the classical calibration estimator given by (10.4.5).

#### $u(\mathbf{x})$ via generalized linear model

<!-- index: Generalized linear model -->What happens if the auxiliary information does not describe the characteristic of interest with linear behavior, as observed in Figure 10.6?

```{r}
#| label: fig-f10-6
#| warning: false
#| message: false
#| fig-height: 5
#| echo: false
#| fig-cap: 'Nonlinear behavior of the characteristic of interest explained by the auxiliary information.'
x <- runif(1000, 1, 3)
y <- rnorm(1000, 4 * exp(x), 2 * x)
dat <- data.frame(x = x, y = y)

ggplot(dat, aes(x = x, y = y)) +
  geom_point(shape = 1) +
  geom_smooth()
```

This is the most important part of the practical development of optimal calibration estimators. In this regard, the user may pause for a moment on the following questions:

- If an auxiliary information characteristic explains the characteristic of interest very well, then calibrating with respect to this auxiliary information would be very convenient. However, this relationship will not always be linear.
- If we want perfect estimates, we should use the characteristic of interest itself for calibration, but since this is absurd, we must use $u(\mathbf{x})$ similar to $y$.

If it is known that the available auxiliary information does not describe the characteristic of interest linearly, the application of the classical calibration estimators motivated by @Dev is called into question. Therefore, if the values of the characteristic of interest are considered realizations of a superpopulation model $\xi$ as in (@eq-modeloopt), which can be described through its first and second moments, then the generalized linear model (GLM), described in detail in @nelder and given by (@eq-glmopt), is clearly relevant. The main feature of the GLM is that the variance of the characteristic of interest depends on the mean $\mu_k$. In addition, in the GLM the characteristic of interest is considered to be related to the auxiliary information variables through the mean $\mu_k$ and a link function $g(\cdot)$ such that
$$
g(\mu_k)=\theta_0+\theta_1x_{k1}+...+\theta_Px_{kP}
$$

Note that the classical linear regression model is a special case of the GLM where $g(\mu_k)=\mu_k$ and $V(\mu_k)=1$. Of course, other forms of the variance function exist, and nonlinear links are also allowed. For example, among the most popular link and variance functions are the logarithmic link given by $g(\mu_k)=\log(\mu_k)$, the Poisson variance function given by $V(\mu_k)=\mu_k$, and the Gamma variance given by $V(\mu_k)=\mu_k^2$.

The GLM is a semiparametric method and requires specifications only for the first and second moments. The link function $\mu_k$ is related to the independent variables, and the variance function describes how the variation in the characteristic of interest is related to the mean.

The coefficients $(\theta_0,\theta_1,...,\theta_k)$ can be estimated, as in our case, using the maximum quasi-likelihood method. In the most general case, the estimator of the vector of population parameters $\hat{\boldsymbol{\theta}}=(\hat{\theta_0},\hat{\theta_1},...,\hat{\theta_P})'$ is the solution of the following equation:
$$
\mathbf{D}'\mathbf{V}^{-1}(\mathbf{y}-\boldsymbol{\mu})=0
$$ {#eq-cap10-4-13}
The preceding equation is simply a generalization of the normal equations in a multiple regression model. Here $\mathbf{y}=(y_1,...,y_n)'$ and $\boldsymbol{\mu}=(\mu_1,...,\mu_n)'$, $\mathbf{V}=diag(V(\mu_1),...,V(\mu_n))$ are the mean and variance structures of the model, respectively, and $\mathbf{D}=\partial\mu/\partial\theta$. The parameters $\theta_p$, $p=1,\ldots,P$, are implicit in (@eq-cap10-4-13). In the simplest case, the classical linear model, we have $\mu_k=\theta_0+\theta_1x_{k1}+...+\theta_Px_{kP}$, $\boldsymbol{\mu}=\mathbf{X}'\boldsymbol{\theta}$, and $\mathbf{D}=\mathbf{X}'$. Then (@eq-cap10-4-13) becomes $\mathbf{X}'\mathbf{V}^{-1}\mathbf{X}\boldsymbol{\theta}=\mathbf{X}'\mathbf{V}^{-1}\mathbf{y}$, which corresponds to the normal equations of multiple regression.

On the other hand, in any other model where the link function is different from the identity, the main difficulty in finding the maximum quasi-likelihood estimator of $\boldsymbol{\theta}$ is that solving (@eq-cap10-4-13) requires iterative procedures.

::: {.callout-tip title="Result"}
Under a GLM superpopulation model, the optimal calibration estimator is given by
$$
\begin{aligned}
\hat{t}_{y,opt}=\hat{t}_{y\pi}+(t_{\hat{\mu}}-\hat{t}_{\hat{\mu}\pi}){\hat{B}_y}
\end{aligned}
$$
with
$$
\hat{B}_y=\frac{\sum_{k\in S}d_kq_k\hat{\mu}_ky_k}{\sum_{k\in S}d_kq_k\hat{\mu}_k^2}
$$
where $\hat{\mu}_k=g^{-1}(\mathbf{x}_k'\boldsymbol{\theta})$ and $g^{-1}(\cdot)$ is the inverse of the link function.

:::
The R statistical software implements the `glm` function, which allows the GLM parameters to be estimated. Suppose that we want to find the maximum quasi-likelihood estimator of $\boldsymbol{\theta}=(\theta_0,\theta_1,...,\theta_P)'$ for the model

$$
\mu_k=\exp(\theta_0+\theta_1x_{k1}), \ \ \ Var_\xi(y_k|\mathbf{x}_k)=\nu(\mu_k)^2=\mu_k^2
$$

Of course, from (10.9.3), the link function is the logarithm. The following lines of code show how to obtain $\hat{\boldsymbol{\theta}}$.

```{r, eval = FALSE}
theta0 <- lm(Y ~ X)
theta1 <- glm(Y ~ X, start = theta0, quasi(var = "mu^2", link = "log"))
```

Note that `theta0` is the least squares estimator and serves as the initial estimator for the iterative process. Analogously, it is possible to create custom code to compute estimates of the parameter vector based on @nelder.

```{r, eval = FALSE}
tol <- 0.000000001
theta0 <- solve(t(X) %*% X, t(X) %*% (Y)) ## initial values
dif <- 1
while (dif >= tol) { ## iteration condition
  mu <- exp(as.vector(X %*% theta0))
  V <- diag(1 / mu)
  theta1 <- theta0 + solve(t(X) %*% X, t(X) %*% V %*% (Y - mu))
  dif <- max(abs(theta1 - theta0))
  theta0 <- theta1
}
```

Of course, the preceding code should match the output produced by R's `glm` procedure.

## Exercises

1. [@TiAr, Exercise 7.1] Using a sampling procedure, the following estimates were obtained for the absolute sizes $\hat{N}_{ij}$ of two subpopulations of interest:

| 80 | 170 | 150 | 400 |
| --- | --- | --- | --- |
| 90 | 80 | 210 | 380 |
| 10 | 80 | 130 | 220 |
| 180 | 330 | 490 | 1000 |

On the other hand, from official sources, the marginal sizes are known with certainty. The true totals for the rows are $(430, 360, 210)$ and the true totals for the columns are $(150, 300, 550)$.

1. Adjust the table to the true marginal population totals using the IPFP algorithm.
1. Adjust the table to the true marginal population totals using the calibration approach with the *raking* method.
1. Explain the differences or similarities between the preceding estimates.

1. [@TiAr, Exercise 7.4] Suppose that the following table was obtained and that the true totals for the rows are $(84, 37, 444, 464)$ and the true totals for the columns are $(49, 859, 11, 10)$.

| 78 | 6 | 0 | 0 | 84 |
| --- | --- | --- | --- | --- |
| 32 | 5 | 0 | 0 | 37 |
| 0 | 0 | 427 | 17 | 444 |
| 0 | 0 | 432 | 32 | 464 |
| 110 | 11 | 859 | 49 | 1029 |

As can be noted, the estimated row totals fully coincide with the true totals. Explain why this table cannot be adjusted using the IPFP algorithm.

1. Consider an agricultural region consisting of $N=2010$ farms, for which a simple random sample of farms of size $n=100$ was selected. In addition, it is known that there are 1580 farms with fewer than 160 hectares (post-stratum 1) and 430 farms with more than 160 hectares (post-stratum 2). The characteristic of interest measured in each farm included in the sample is the cereal-cultivated area on each farm. If the realized sample had $n_1=70$, $n_2=30$, $\bar{y}_1=19.4$, and $\bar{y}_2=51.63$, estimate, using the calibration technique, the population mean of the cereal-cultivated area in the agricultural region and report the estimated coefficient of variation.

1. Consider a Poisson sampling design with unequal inclusion probabilities $\pi_k$, $k\in U$. Suppose that the population total $t_y$ is of interest. Construct a calibration estimator using a single auxiliary information characteristic $x_k=1$ and $q_k=1$, for all $k\in U$, using the following pseudo-distance (parameterized by $\alpha$):

$$
G(x)=
\begin{cases}
\frac{1}{\alpha(\alpha-1)}(x^{\alpha}+(\alpha-1)-\alpha x), & \text{if $\alpha \in \mathbb{R}-\{0,1\}$} \\
x\ln(x)+1-x, & \text{if $\alpha=1$} \\
\ln(1/x)-1+x, & \text{if $\alpha=0$}
\end{cases}
$$

1. Write the calibration equations.
1. Obtain the function $g(x)$ for the three cases of $\alpha$.
1. Show that the function $F(u)$ is fixed and equal to $N/\hat{N}$.
1. Deduce the calibration weights.
1. Obtain the resulting calibration estimator. What form does the resulting estimator have?

1. Suppose that the information from Exercise 8.7 is the result of a Poisson sampling plan with inclusion probability $\pi_k=n (x_k/t_x)$. Using the results of the previous exercise and assuming that $x_k=1$ and $q_k=1$ for all $k\in U$, obtain a calibration estimate for the total number of inhabitants in the municipality, the number of automobiles in the municipality, and the number of military personnel in the municipality. Obtain the corresponding estimated coefficients of variation.

1. Support or refute the following statements.
1. Calibration estimators induced by the chi-square distance fully coincide with general regression estimators.
1. The quantity $q_k$ is constant for all individuals under the entropy distance.
1. Under the inverse chi-square distance, minimizing the distance with respect to the calibration restrictions always leads to weights $w_k$ equal to the inverse of the inclusion probability of the $k$-th element.

::: {.citation-box}
### How to cite this book  {-}

Gutiérrez, A. (2027). *Sampling strategies: Survey design and parameter estimation*. Unpublished version online.
:::
